---
title: "codeprez"
output: html_notebook
---

# Packages installation

```{r eval=FALSE, paged.print=TRUE}
install.packages(c("devtools", "tidyverse", "kableExtra", "gt", "glue"))
devtools::install_github("waldronlab/bugSigSimple")
devtools::install_github("waldronlab/BugSigDBStats")
devtools::install_github("waldronlab/bugsigdbr")
```

# Data import, recoding, and subset
```{r results = FALSE}
library(bugSigSimple)
library(BugSigDBStats)
dat <- bugsigdbr::importBugSigDB(cache = FALSE) 
dim(dat)
names(dat)
```

```{r results = FALSE}
library(tidyverse)
condsnew <- c("COVID-19")
covid_all <- subsetByCondition(dat, condsnew) %>%
  mutate(studyexp = paste(Study, Experiment, sep = "_")) %>%
  mutate(
    site = recode(`Body site`,
      "feces" = "Gut",
      "rectal" = "Gut",
      "nasopharynx" = "aURT",
      "oropharynx" = "aURT",
      "nasopharynx,oropharynx" = "aURT",
      "tongue" = "aURT"
    )
  ) %>%
  mutate(comparison1 = paste(`Group 0 name`, `Group 1 name`, sep = " vs ")) %>%
select(-c(Curator, `Curated date`, `Revision editor`, `LDA Score above`, `Matched on`, `Significance threshold`, `Confounders controlled for`))
View(covid_all)

library(stringr)
regexp <- "[[:digit:]]+"
studyno <- str_extract(covid_all$Study, regexp)
expno <- str_extract(covid_all$Experiment, regexp)
signo <- str_extract(covid_all$`Signature page name`, regexp)
cbind(covid_all, studyno, expno, signo)

covid_all2 <- covid_all %>%
  mutate(ID = gsub(" ", "", paste("bsdb:", `studyno`, "/", `expno`, "/", `signo`))) %>%
  mutate()
  
View(covid_all2)
```

##subset data by bodysite
```{r}
gut_sigs <- filter(covid_all2, 
                           site == "Gut") %>%
    drop_na(Source)
URT_sigs <- filter(covid_all2, 
                           site == "aURT") %>%
  drop_na(Source)
naso_sigs <- filter(covid_all2, 
                           `Body site` == c("nasopharynx", "nasopharynx,oropharynx"))
oro_sigs <- filter(covid_all2, 
                           `Body site` == c("oropharynx", "nasopharynx,oropharynx"))
```


## Binomial Testing

In this table, the Binomial Test p-value corresponds to the null hypothesis

H0: the proportion of signatures in which the taxon is reported increased or decreased, relative to the total number of signatures in which it is reported, is equal to 0.5

##binomial testing for all gut signatures
```{r results = TRUE}
library(kableExtra)
kableExtra::kbl(bugSigSimple::createTaxonTable(gut_sigs))
```

## binomial testing for case-control gut signatures

Filter for case-control comparisons [healthy controls vs COVID patients at any stage of infection]
```{r results = FALSE}
library(dplyr)
healthy <- grepl(".*(healthy.*COVID|COVID.*healthy).*", gut_sigs$comparison1, ignore.case = TRUE)
cc_gut_sigs1 <- gut_sigs %>%
      filter(healthy == TRUE) 

recovered <- grepl("Recovered", gut_sigs$`Group 1 name`, ignore.case = TRUE)
rec_gut_sigs <- gut_sigs %>%
  filter(recovered == TRUE)
View(rec_gut_sigs)
```

```{r results = TRUE}
library(bugSigSimple)
cc_gut_sigs1_taxontable <- bugSigSimple::createTaxonTable(cc_gut_sigs1) %>% 
  mutate(FDR =  p.adjust(p = `Binomial Test pval`, method="fdr")) %>%
  relocate(FDR, .after = `Binomial Test pval`)
kableExtra::kbl(cc_gut_sigs1_taxontable)
```

##add categorical variables for studies of interest (case/control, recovered pts, 2 sick groups) in each subset dataframe
```{r results = FALSE}
library(dplyr)
healthynaso <- grepl(".*(control.*COVID|COVID.*control).*", naso_sigs$comparison1, ignore.case = TRUE)
recovnaso <- grepl("recovered*", naso_sigs$comparison1, ignore.case = TRUE)
sicknaso <- grepl(".*(patients.*patients|patients.*patients).*", naso_sigs$comparison1, ignore.case = TRUE)
naso_sigs1 <- naso_sigs %>%
      mutate(casecont = ifelse(healthynaso, 1, 0)) %>%
   mutate(recov = ifelse(recovnaso, 1, 0)) %>%
mutate(sick2 = ifelse(sicknaso, 1, 0))

healthyoro <- grepl(".*(healthy.*COVID|COVID.*healthy).*", oro_sigs$comparison1, ignore.case = TRUE)
recovoro <- grepl("recovered*", oro_sigs$comparison1, ignore.case = TRUE)
sickoro <- grepl(".*(patients.*patients|patients.*patients).*", oro_sigs$comparison1, ignore.case = TRUE)
oro_sigs1 <- oro_sigs %>%
     mutate(casecont = ifelse(healthyoro, 1, 0)) %>%
   mutate(recov = ifelse(recovoro, 1, 0)) %>%
mutate(sick2 = ifelse(sickoro, 1, 0))

healthygut <- grepl(".*(healthy.*COVID|COVID.*healthy).*", gut_sigs$comparison1, ignore.case = TRUE)
recovgut <- grepl("recovered*", gut_sigs$comparison1, ignore.case = TRUE)
sickgut <- grepl(".*(patients.*patients|patients.*patients).*", gut_sigs$comparison1, ignore.case = TRUE)
gut_sigs1 <- gut_sigs %>%
      mutate(casecont = ifelse(healthygut, 1, 0)) %>%
  mutate(recov = ifelse(recovgut, 1, 0)) %>%
   mutate(sick2 = ifelse(sickgut, 1, 0))

URT_sigs1 <- rbind(naso_sigs1, oro_sigs1)

all_sigs1 <- rbind(naso_sigs1, oro_sigs1, gut_sigs1)
```

# Cluster analysis

First calculate pairwise overlaps for all signatures of length > 1:
## OVERALL CLUSTER ANALYSIS - all signatures
```{r}
allsigs <- bugsigdbr::getSignatures(all_sigs1, tax.id.type = "taxname")
allsigs <- allsigs[sapply(allsigs, length) > 1] #require length > 1
dim(allsigs)
mydists_all <- BugSigDBStats::calcPairwiseOverlaps(allsigs)
dim(mydists_all)
View(allsigs)
```

```{r}
allsignames <- unique(c(mydists_all$name1, mydists_all$name2))
jmat_all <- matrix(NA, nrow=length(allsignames), ncol=length(allsignames), dimnames=list(allsignames, allsignames))
diag(jmat_all) <- 1
for (i in 1:nrow(mydists_all)){
  jmat_all[mydists_all[i, "name2"], mydists_all[i, "name1"]] <- mydists_all[i, "jaccard"]
  jmat_all[mydists_all[i, "name1"], mydists_all[i, "name2"]] <- mydists_all[i, "jaccard"]
}
View(jmat_all)
```

```{r, fig.height=60, fig.width=20, results = TRUE}
hc_all <- hclust(as.dist(jmat_all))
plot(hc_all)
```

```{r}
library(dplyr)
clustsall <- sort(cutree(hc_all, h = 0.05))
lapply(unique(clustsall), function(i) names(clustsall)[clustsall == i])

clustnames<- as.vector(names(clustsall))
clustval <- as.vector(clustsall)
cn <- data.frame(clustnames, clustval) %>%
  mutate(ID = substr(clustnames, 1, 12))

allclust <- merge(all_sigs1, cn, by = "ID") %>%
  mutate_if(is.character, as.factor) %>%
  mutate(as.factor(casecont)) %>%
  mutate(as.factor(clustval))
str(allclust)
```

```{r warning = FALSE, results = TRUE}
match(c("Body site", "Location of subjects", "Sequencing type", "Sequencing platform", "casecont", "clustval", "recov",  "sick2"), names(allclust))

chisqall <- lapply(allclust[,c(13,15,23,25,43,44,45)], function(x){ chisq.test(allclust[,49], x)}
               )
do.call(rbind,chisqall)[,c(1,3)]
```

###silhouette plot
```{r}
library(cluster)
dev.new(width = 10, height = 5, unit = "in")
plot(silhouette(cutree(hc_all, h = 0.05),jmat_all))
```

##gut cluster analysis
```{r}
library(dplyr)
sigs.gut <- bugsigdbr::getSignatures(gut_sigs, tax.id.type = "taxname")
sigs.gut <- sigs.gut[sapply(sigs.gut, length) > 1] #require length > 1
dim(sigs.gut)
mydists_gut <- BugSigDBStats::calcPairwiseOverlaps(sigs.gut)

gut2 <- gut_sigs1 %>%
  select("studyexp","Body site", "Location of subjects", "Sequencing type", "Sequencing platform", "comparison1", "ID", "casecont", "recov", "sick2") %>%
  mutate(sigs = bugsigdbr::getSignatures(gut_sigs, tax.id.type = "taxname")) 

sigs.gut2 <- gut2[sapply(gut2$sigs, length) > 1,] %>%
  mutate(location = as.factor(`Location of subjects`)) %>%
           mutate(seqtype = as.factor(`Sequencing type`)) %>%
                    mutate(seqplat = as.factor(`Sequencing platform`)) %>%
  select(-c("Location of subjects", "Body site", "Sequencing type","Sequencing platform"))
```

```{r}
gutsignames <- unique(c(mydists_gut$name1, mydists_gut$name2))
jmat_gut <- matrix(NA, nrow=length(gutsignames), ncol=length(gutsignames), dimnames=list(gutsignames, gutsignames))
diag(jmat_gut) <- 1
for (i in 1:nrow(mydists_gut)){
  jmat_gut[mydists_gut[i, "name2"], mydists_gut[i, "name1"]] <- mydists_gut[i, "jaccard"]
  jmat_gut[mydists_gut[i, "name1"], mydists_gut[i, "name2"]] <- mydists_gut[i, "jaccard"]
}
View(jmat_gut)
```

```{r, fig.height=60, fig.width=20, results = TRUE}
hc_gut <- hclust(as.dist(jmat_gut))
plot(hc_gut)
```

###silhouette plot
```{r results = TRUE}
library(cluster)
dev.new(width = 10, height = 5, unit = "in")
plot(silhouette(cutree(hc_gut, h = 0.05),jmat_gut))
```

```{r}
clustsgut <- sort(cutree(hc_gut, h = 0.05))
lapply(unique(clustsgut), function(i) names(clustsgut)[clustsgut == i])

clustnames<- as.vector(names(clustsgut))
clustval <- as.vector(clustsgut)
cn <- data.frame(clustnames, clustval) %>%
  mutate(ID = substr(clustnames, 1, 12))

gutclust <- merge(gut2, cn, by = "ID") %>%
  select(-c("sigs")) %>%
  mutate_if(is.character, as.factor) %>%
  mutate(as.factor(casecont)) %>%
  mutate(as.factor(clustval))
str(gutclust)
```

##are any of the following variables (seq type, seq platform, case/control vs not, etc) associated with cluster group?
```{r, warning=FALSE}
match(c("casecont", "Location of subjects", "Sequencing type", "Sequencing platform", "clustval", "recov",  "sick2"), names(gutclust))
chisqgut <- lapply(gutclust[,c(4,5,6,8,9,10)], function(x){ chisq.test(gutclust[,12], x)} 
                    )
do.call(rbind,chisqgut)[,c(1,3)] 
```


##SEMANTIC SIMILARITY ANALYSIS
```{r}
library(ontologyIndex)
onto <- ontologyIndex::get_ontology("http://purl.obolibrary.org/obo/ncbitaxon.obo")
```

```{r heinzcache, echo = FALSE} 
onto <- BugSigDBStats:::.getResourceFromCache("ncbi.onto")
if(is.null(onto))
{
    onto <- ontologyIndex::get_ontology("http://purl.obolibrary.org/obo/ncbitaxon.obo")
    BugSigDBStats:::.cacheResource(onto, "ncbi.onto")
} 
```

```{r}
onto
head(onto$id)
```

We add the corresponding ID prefix:

```{r}
sigs <- bugsigdbr::getSignatures(gut_sigs, tax.id.type = "ncbi")
sigs <- lapply(sigs, function(s) paste0("NCBITaxon:", s))
```            

We remove taxa that are not in the NCBI Taxonomy:

```{r}
utax <- unique(unlist(sigs))
nt <- utax[!(utax %in% onto$id)]
sigs <- lapply(sigs, function(s) setdiff(s, nt))
```

Now, we compute pairwise semantic similarity for all signatures:

```{r}
library(ontologySimilarity)
sim.mat <- ontologySimilarity::get_sim_grid(ontology = onto, term_sets = sigs)
sim.mat[1:5,1:5]
```

Given the matrix of pairwise semantic similarity between signatures, we can also
compute the semantic similarity of a group of signatures by eg. taking the average
similarity between all pairs of signatures.

Here, we compute the semantic similarity of all COVID-19 gut signatures in
the database for which abundance is increased in the cases. 

```{r}
ind <- !is.na(gut_sigs[["Condition"]]) &
       gut_sigs[["Condition"]] == "COVID-19" & 
       gut_sigs[["Body site"]] == "feces" & 
       gut_sigs[["Abundance in Group 1"]] == "increased"
ontologySimilarity::get_sim(sim.mat, group = which(ind))
```

Furthermore, we can compute a p-value for assessing the statistical
significance of the similiarity of a group of signatures. The p-value is calculated
by random sampling of groups ofthe same size as ‘group’, and calculating how many
random groups have at least as great group similarity than does ‘group’.

```{r}
ontologySimilarity::get_sim_p(sim.mat, group = which(ind))
```
